---
title: "Usenet Text Analysis"
subtitle: "Text Mining & Topic Modeling"
author: "Zach Palmer"
date: "2025-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set Up & Load Data

```{r load-packages, message=FALSE, warning=FALSE}
library(stringr)
library(tidyverse)
```

```{r}
training_folder = "data/20news-bydate/20news-bydate-train/"

read_folder = function(folder) {
  tibble(file = dir(folder, full.names = TRUE)) %>%
    mutate(text = map(file, read_lines)) %>%
    transmute(id = basename(file), text) %>%
    unnest(text)
}

raw_text = tibble(folder = dir(training_folder, full.names = TRUE)) %>%
  mutate(folder_out = map(folder, read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), id, text)
```

### Quick Look Into the Data

```{r}
raw_text %>% 
  glimpse()
```

```{r}
raw_text %>%
  head(10)
```

**Note:** These first 10 rows are a perfect example of something noted later on: each message begins with a header that contains field such as “from:” or “in_reply_to:” that describe the message.

```{r, fig.height=6, fig.width=8}
raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  mutate(newsgroup = reorder(newsgroup, -messages)) %>%
  ggplot(aes(messages, newsgroup)) +
  geom_col(fill = "dodgerblue") +
  theme_bw() +
  labs(title = "Number of Messages by Usenet Newsgroup", 
       x = "Number of Messages", y = "")
```

**Note:** Each Usenet newsgroup is named hierarchically (i.e. tree structure), each starting with a main topic, such as “talk”, “sci”, or “rec”, and each subsequent topic separated by a period further segmenting the newsgroup into a distinct discussion area. 

### Pre-processing Text

"Each message has some structure and extra text that we don’t want to include in our analysis. For example, every message has a header, containing field such as “from:” or “in_reply_to:” that describe the message. Some also have automated email signatures, which occur after a line like `--`. 

```{r}
# must occur after the first occurrence of an empty line,
# and before the first occurrence of a line starting with --
cleaned_text = raw_text %>%
  group_by(newsgroup, id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(text, "^--")) == 0) %>%
  ungroup()
```

"Many lines also have nested text representing quotes from other users, typically starting with a line like “so-and-so writes…” These can be removed with a few regular expressions."

```{r}

```





